From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction

### 2015-09-30 paper.tex
Improved the overall coherence of the paper by better relating attention tracking to withmeness

This article is organized in three sections: we first present a novel method for
on-line estimation of the focus of attention based on fast 6D head pose
estimation. We then validate this approach with a real-world field experiment
with children. We finally show how to compute ``with-me-ness'' over the course
of the interaction to provide a new \emph{in-the-moment} measurement of the
quality of interaction, and compare levels computed by the robot with a ground
truth based on manual post-hoc annotations.

@@

The main results are reported in Table~\ref{tab:results}
%and Figure~\ref{fig:cosinematrix} (cosine similarity matrix);
Figure~\ref{fig:realExpected} further gives a concrete picture of the
ground-truth \vs computed attentional targets for subject 4 (the subject with
the \emph{least} successful tracking).

@@

@@

\section{With-me-ness}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{with-me-ness}
    \caption{\small \textbf{With-me-ness}. Evolution of the level of
        \emph{with-me-ness} over the whole $\approx$17min long interaction of
        subject 2. The top chart is similar to Figure~\ref{fig:realExpected}
        with the \emph{expected} attentional targets added in green. The bottom
        diagram represents the instantaneous level of with-me-ness over a
        sliding window of 30 seconds. The blue line is the with-me-ness as estimated
        by the robot, the orange line is the with-me-ness computed from manually
        annotated attentional targets.  Pearson's correlation between both
        series for this subject: $r(973)=0.58, p < 0.001$.}

@@

\section{Discussion and Conclusion}

We have presented how a robot can effectively assess in real-time with a regular
camera, the focus of attention of its interactant, and how we can combine it
with the robot's {\it a priori} knowledge about the interaction to build a
metric of \emph{with-me-ness} over the course of the interaction.

@@

>>>>

The following sections expose the full methodology, starting with the estimation
of attention: we present in the next section a novel method for on-line estimation
of the focus of attention based on fast 6D head pose estimation. We validate
this technique in section~\ref{sec:expe} with a real-world field study involving
children. Section~\ref{sec:withmeness} formally introduce the concept of
\emph{with-me-ness}. We present how to compute it over the course of the
interaction to eventually build a new \emph{in-the-moment} measurement of the
quality of interaction. We finally validate and discuss this metric by comparing
it to manual post-hoc annotations of the video-recordings of the interaction.

@@

The main results are reported in Table~\ref{tab:results}
%and Figure~\ref{fig:cosinematrix} (cosine similarity matrix);
Figure~\ref{fig:realExpected} further gives a concrete picture of the
ground-truth \vs computed attentional targets for subject~4 (the subject with
the \emph{least} successful tracking).

@@

The next section builds upon this technique for real-time estimation of
the focus of attention: by comparing the focus of attention with the set of
attentional targets {\it a priori} expected by the robot, we can estimate to
what extent the user is ``with'' the robot.

@@

\section{With-me-ness}
\label{sec:withmeness}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{with-me-ness}
    \caption{\small \textbf{With-me-ness}. Evolution of the level of
        \emph{with-me-ness} over the whole $\approx$17min long interaction of
        subject~2. The top chart is similar to Figure~\ref{fig:realExpected}
        with the \emph{expected} attentional targets added in green. The bottom
        diagram represents the instantaneous level of with-me-ness over a
        sliding window of 30 seconds. The blue line is the with-me-ness as estimated
        by the robot, the orange line is the with-me-ness computed from manually
        annotated attentional targets.  Pearson's correlation between both
        series for this subject: $r(973)=0.58, p < 0.001$.}

@@

\section{Discussion}

@@

\section{Conclusion}

We have presented how a robot can effectively assess in real-time with a regular
camera, the focus of attention of its interactants, and how we can combine it
with the robot's {\it a priori} knowledge about the interaction to build a
metric of \emph{with-me-ness} over the course of the interaction.

The experimental validation has been conducted with six children in face-to-face
interaction with an autonomous robot, over a total duration of about 2 hours. It
shows that {\it 1)} most of the time, we are able to estimate the head pose of
these children; {\it 2)} based on these head poses only, the instantaneous focus
of attention as computed by the robot does reach a good level of accuracy, with
however one inaccurate outlier out of the six participant; {\it 3)} the robot is
able to compute in real-time a level of \emph{with-me-ness} that correlates
strongly with the ground-truth.

The accuracy of the attention estimation could be improved, first by adding
heuristics to detect and ignore erroneous/inaccurate face detections, second by
implementing pupil tracking on top of head pose estimation. However, our results
show that the accuracy levels that we reach already support the reliable
computation of a metric measuring a high-level social construct, the
\emph{with-me-ness}, which we argued should be used as a well-defined, reliable
precursor of engagement in building adaptive robot behaviors.
